---
title: "Re-analysing the data from @moffatt_inner_2020: A textbook illustration of the absence of evidence fallacy"
shorttitle: "The absence of evidence fallacy"
author: 
  - name: Ladislas Nalborczyk
    affiliation: 1
    corresponding: yes
    address: GIPSA-lab, CNRS, Univ. Grenoble Alpes, 11 Rue des Mathématiques, 38400 Saint-Martin-d'Hères, France
    email: ladislas.nalborczyk@gipsa-lab.fr
affiliation:
  - id: 1
    institution: Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France
abstract: "@moffatt_inner_2020 reported the results of an experiment (N = 26 in the final sample) comparing the facial (surface) electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on the absence of significant difference in the perioral muscular activity between the rumination and distraction conditions, @moffatt_inner_2020 concluded that *self-reported* inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is hasty and based on waggly evidence. Indeed, concluding on the absence of an effect based on a low-powered non-significant p-value is strongly problematic/uninformative. Moreover, the relation between self-reports and physiological measures was not *directly* assessed, but only indirectly inferred from differences (or absence thereof) in group means. Given the ample inter-individual variability in these measures (as suggested by our reanalysis), we think inferring the individual-level relation between self-reports and physiological measures from group means is inappropriate. Given these limitations, we conclude that it is unclear whether the target article adds to the current/extent knowledge and we suggest ways forward, both from a theoretical and from a methodological perspective. Complete source code, reproducible analyses, and figures are available at https://github.com/lnalborczyk/inner_experience_EMG."
keywords: NHST, Bayesian, fallacy, reanalysis, inner speech, rumination, electromyography
csl: csl/apa7.csl
classoption: man
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false
bibliography: [bib/library.bib, bib/r-references.bib]
nocite: | 
  @R-base, @R-knitr, @R-rmarkdown, @R-tidyverse, @R-papaja, @R-wordcountaddin,
  @R-here
output:
  papaja::apa6_pdf:
    keep_tex: true
    highlight: tango
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(wordcountaddin)
library(BayesFactor)
library(ggbeeswarm)
library(rmarkdown)
library(tidyverse)
library(papaja)
library(readxl)
library(knitr)
library(brms)
library(here)
library(pwr)

# setting seed for reproducibility
# set.seed(666)

# setting up knitr options
knitr::opts_chunk$set(
  cache = TRUE, eval = TRUE, echo = FALSE,
  warning = FALSE, message = FALSE,
  fig.align = "center", dev = "pdf",
  out.width = "100%", fig.width = 8,
  fig.pos = "!htb"
  )

# counting words in the current .Rmd file
wordcount <- word_count(knitr::current_input() )
```

Wordcount (excluding abstract, references, tables, and figures): `r wordcount`

\newpage

<!-- NB: You can add comments using these tags -->

# Introduction

The activity of silently talking to oneself or "inner speech" is a foundational ability... despite its multiple adaptive functions in everyday life, inner speech can go awry and leads to sustained negative... These inner speech "dysfunctions" [for reviews, see @alderson-day_inner_2015; @perrone-bertolotti_what_2014; @loevenbruck_cognitive_2018]...

Given the predominantly verbal nature of rumination [], we previously proposed to study rumination as other forms of inner speech have been studied in the past, namely using surface electromyography and motor interference protocols [e.g., @nalborczyk_orofacial_2017; @nalborczyk_understanding_2019; @nalborczyk_can_2020; @nalborczyk_articulatory_2020]...

We have previously shown that... However, it was unclear... therefore, the extension of our study by @moffatt_inner_2020, which consisted by including a distraction control group, is more than welcome... However...

The main conclusion from @moffatt_inner_2020 is that inner experience between induced rumination and distraction differs "without a change in electromyographic correlates of inner speech". In other words, their conclusion is that inner experience is unrelated (or loosely related) to the electromyographic correlates of inner speech, which are thought to be represented mostly by the EMG amplitude recorded over the OOI and OOS muscles. However, for this in-sample observation to be of interest in an out-of-sample context (i.e., to be informative of other non-observed individuals, or said otherwise, to bring information about the population), this absence of difference has to be based on sufficiently powered sample size (given the target effect size) as well as on reliable measures. Moreover, a simple visual exploration of the data reveals important variability between individuals in the main effect of interest. That is, some participants had higher perioral (OOS and OOI) muscular activity in the rumination condition than in the distraction condition, and some other participants showed the reverse pattern. This suggests unexplored variation in the determinants of this effects (e.g., the content of the inner experience). Indeed, the relation between the inner experience and the physiological correlates of inner speech production was only inferred from group means. However, given the previous point, this appears highly problematic. We explore each of these limitations and suggests ways forward in the following section.

# Exploring the data

```{r importing_data}
# importing data
df <- read_excel(path = "data/RUM_master_data.xlsx", sheet = 1)

# importing codebook
codebook <- read_excel(path = "data/RUM_master_data.xlsx", sheet = 2)

# data reshaping (from wide to long)
# NB: 26 participants in the final dataset...

# Order in which rumination and distraction inductions were completed
# 1 = rumination induction completed first
# 2 = distraction induction completed first

df2 <- df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # extracts muscle name
    mutate(muscle = substr(muscle, 1, 3) ) %>%
    # from long to wide
    pivot_wider(names_from = muscle, values_from = value)
```

```{r general, fig.cap = "Average log-EMG amplitude by muscle and condition. The black dots and intervals represent the by-group average and 95\\% confidence interval (N = 26). The horizontal white line in the violin plot represents the median. The grey dots represent the individual-level average natural logarithm of the EMG amplitude by muscle and condition."}
# position for groups
pd <- position_dodge(0.8)

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.15,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun = mean,
        geom = "line", size = 1,
        # aes(group = interaction(muscle, order) ),
        aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    # axis labels
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

...

```{r model, results = "hide"}
# priors for the varying-effects model
priors_varying <- c(
    prior(normal(0, 10), class = Intercept, resp = "FRO"),
    prior(normal(0, 1), class = b, resp = "FRO"),
    prior(exponential(1), class = sigma, resp = "FRO"),
    prior(exponential(1), class = sd, resp = "FRO"),
    prior(normal(0, 10), class = Intercept, resp = "OOS"),
    prior(normal(0, 1), class = b, resp = "OOS"),
    prior(exponential(1), class = sigma, resp = "OOS"),
    prior(exponential(1), class = sd, resp = "OOS"),
    prior(normal(0, 10), class = Intercept, resp = "OOI"),
    prior(normal(0, 1), class = b, resp = "OOI"),
    prior(exponential(1), class = sigma, resp = "OOI"),
    prior(exponential(1), class = sd, resp = "OOI")
    )

# fitting the model
varying_effects <- brm(
    mvbind(FRO, OOS, OOI) ~ 1 + condition + (1 | ID),
    family = gaussian(),
    prior = priors_varying,
    data = df2,
    chains = 4, cores = detectCores(),
    warmup = 2000, iter = 10000,
    control = list(adapt_delta = 0.95),
    sample_prior = TRUE,
    file = here("models/varying_effects")
    )
```

Posterior distribution of the difference between the distraction and rumination conditions...

## Concluding on the null from low-powered studies is uninformative

There is an infamous tradition of running uninformative null-hypothesis significance tests in Psychology [e.g., @harlow_problem_1997; @meehl_appraising_1990-1; @meehl_theoretical_1978; @meehl_theory-testing_1967; @meehl_why_1990]. By "uninformative", we mean that some null-hypothesis significance tests are often *not* diagnostic with regards to the substantive question of interest...

As highlighted by many authors [e.g., @pollard_probability_1987; @rouder_is_2016], concluding on an absence of difference based on not obtaining evidence for the difference is the continuous extension of the logical fallacy of the... The argument from ignorance, such as "Science has found no proof of intelligent life nearby us in space, therefore intelligent life does not exist nearby us in space."... the absence of evidence fallacy or fallacy of acceptance...

This problem is tackled in modern usages of null-hypothesis significance test by ensuring that the test has good *severity* [e.g., @mayo_severe_2006; @mayo_statistical_2018]. In general terms, we have evidence for a claim to the extent that it survives a stringent scrutiny, that is if it survives *severe tests*. In other words, some claim (e.g., $\theta = 0$) is said to be *severely tested*) if it had great chances of being falsified, was the claim false. More formally, we can define $\text{SEV}(T, x0, H)$, the severity with which claim $H$ passes test $T$ with outcome $x0$, and $\text{SEV}(\mu > \mu_{1}) = \Pr(d(X) \leq d(x0); \mu = \mu_{1})$ [@mayo_severe_2006; @mayo_statistical_2018]...To put it simply... https://www.analytics-toolkit.com/glossary/severity/...

Anticipating the critics on the power of their study (a critic that was probably raised during peer review), @moffatt_inner_2020 report the results of a (possibly ran a posteriori) power analysis using the effect size reported in @nalborczyk_orofacial_2017 of $d = 0.72$, which is highly optimistic estimate of the substantive effect of interest in the target article (i.e., the difference in EMG amplitude between the rumination and distraction conditions) as this effects represents the standardised mean difference *between a rest period and a rumination one* [@nalborczyk_orofacial_2017]...

```{r power, echo = TRUE}
# How many participants do we need for a target statistical power of 0.8?
library(pwr)
pwr.t.test(
  d = 0.72, sig.level = 0.05, power = 0.8,
  type = "one.sample", alternative = "two.sided"
  )
```

We suggest the (a priori) power of the study ran by @moffatt_inner_2020 was much lower than suggested by the authors. Indeed, we may speculate that the effect (i.e., the standardised mean difference in EMG amplitude) between the rumination and distraction conditions may be much weaker than the effect (i.e., the standardised mean difference in EMG amplitude) between the rumination and the rest conditions. If we assume that the former is half the size of the latter [which seems reasonable given the high inter-individual variability in such effects, cf. the next section but also @nalborczyk_can_2020], therefore the a priori power of the main statistical test from @moffatt_inner_2020 was around $0.44$, meaning that they had less than 1 chance out of 2 to find a significant effect, given that the effect in the population was actually $0.36$. Because this is less than the chance of obtaining a head in a coin flip, we feel these resources may have been better invested.

```{r actual_power, echo = TRUE}
# A priori power for n = 26 (per condition) and d = 0.36
pwr.t.test(
  n = 26, d = 0.72 / 2, sig.level = 0.05,
  type = "one.sample", alternative = "two.sided"
  )
```

Anticipating the legitimate critique that the absence of a significant difference is not *necessarily* "significant" evidence for the absence of an effect, @moffatt_inner_2020 reported the following Bayes factor (BF) analysis:

> "[...] therefore it is possible that the sample size of the present study lacked sufficient power to detect the effect of rumination on muscle activity. In order to test this, a Bayesian paired samples t-test was conducted for the peak log values of muscle activity between the rumination and distraction conditions. This revealed strong evidence in favour of the alternative hypothesis for the FRO muscle ($B_{10} = 18.79$), and moderate evidence in favour of the null hypothesis for the OOS ($B_{10} = 0.232$) and OOI ($B_{10} = 0.278$) muscles, according to current guidelines for interpreting Bayes factors [43]."

<!--

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, computing a BF (i.e., comparing two models) does not solve *at all* the problem of low power (we discuss this issue in more depth below). Second, few details are given with regards to the exact models that were compared. Third, and most importantly, the BFs indicate moderate evidence in favour of the null for the OOI and OOS muscles. More precisely, these BFs indicated that the (observed) data are $1 / 0.232 \approx 4.31$ times more likely under the null than under the alternative hypothesis for the OOS and $1 / 0.278 \approx 3.6$ times more likely under the null than under the alternative hypothesis for the OOI. In other words, the evidence is favour of the null is relatively weak and sensitivity analyses (i.e., reporting the BF with varying prior scales for the alternative hypothesis) may unsurprisingly results in various BFs. For instance... 

-->

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, computing a BF (i.e., comparing two models) does not solve *at all* the problem of low power...

```{r sensitivity, echo = FALSE, eval = FALSE}
library(BayesFactor)
# rscale = sqrt(2) / 2
ttestBF(x = df2$OOI[df2$condition == "RUM"], y = df2$OOI[df2$condition == "DIS"], paired = TRUE, rscale = "medium")
# rscale = 1
ttestBF(x = df2$OOI[df2$condition == "RUM"], y = df2$OOI[df2$condition == "DIS"], paired = TRUE, rscale = "wide")
# rscale = sqrt(2)
ttestBF(x = df2$OOI[df2$condition == "RUM"], y = df2$OOI[df2$condition == "DIS"], paired = TRUE, rscale = "ultrawide")
```

We first fitted a multivariate Bayesian regression model with varying-intercepts (by participant) and weakly informative priors on these data using the `brms` package [@R-brms]. From there, we i) generated new datasets from the posterior predictive distribution and ii) we computed the BF in favour of the alternative hypothesis ($\text{BF}_{10}$) using the `BayesFactor` package [@R-BayesFactor]. We used a "medium" prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with $1000$ simulations (i.e., 1000 simulated datasets) for each sample size.

```{r simulated-power, fig.cap = "Average natural logarithm of the Bayes factor in favour of the alternative hypothesis, along with its standard error, computed over 1000 datasets of increasing size simulated from the posterior predictive distribution of the varying-intercept multivariate Bayesian regresion model. A log-BF belows 0 represents evidence for the null hypothesis (relative to the alternative) and a log-BF above 0 represents evidence for the alternative hypothesis (relative to the null)."}
# running the simulations
# source("code/bf_simulation.R")

# loading the results
load(file = "results/overall_results.Rda")

# simulations parameters
nsims <- 1e3
sample_size <- seq.int(from = 20, to = 200, by = 10)

# plotting the results
overall_results %>%
    na.omit() %>%
    filter_all(any_vars(. != 0) ) %>%
    pivot_longer(cols = bf_fro:bf_oos, names_to = "bf_type") %>%
    mutate(value = log(value) ) %>%
    mutate(bf_type = factor(bf_type, labels = c("FRO", "OOI", "OOS") ) ) %>%
    group_by(nobs, bf_type) %>%
    summarise(
        across(
            .cols = value,
            .fns = list(mean = mean, se = ~sd(.x) / sqrt(nsims) )
            )
        ) %>%
    ungroup() %>% 
    ggplot(aes(x = nobs, y = value_mean, colour = bf_type, fill = bf_type) ) +
    geom_hline(yintercept = 0, lty = 3) +
    # geom_raster(fill = value_mean) +
    geom_ribbon(
        aes(
          x = nobs, ymin = value_mean - 1.96 * value_se,
          ymax = value_mean + 1.96 * value_se, colour = NULL
          ),
        alpha = 0.5, show.legend = FALSE
        ) + 
    geom_line(show.legend = FALSE) +
    geom_point(show.legend = FALSE) +
    facet_wrap(~bf_type, scales = "free") +
    scale_x_continuous(breaks = unique(sample_size)[c(TRUE, FALSE)]) +
    theme_bw(base_size = 10) +
    labs(x = "Number of participants", y = "Natural logarithm of the Bayes factor") +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

As shown in Figure \@ref(fig:simulated-power), the natural logarithm of the BF in favour of the alternative hypothesis is growing proportionally with the sample size. More precisely, whereas low sample sizes (i.e., below 80) support the null hypothesis, adequately-powered sample sizes support the alternative hypothesis for all three facial muscles for sample sizes larger than 80 participants. For instance, the average $\text{BF}_{10}$ computed for the OOI muscle with a sample size of 160 participants is of $\exp(2.18) \approx 8.85$, indicating that these data are approximately $8.85$ times more likely under the alternative hypothesis than under the null hypothesis.^[Alternatively, the BF can be interpreted as an updating factor, from prior odds to posterior odds.]

We should keep in mind some limitations of this analysis, which uses simulated datasets form the posterior predictive distribution estimated on the data collected by @moffatt_inner_2020. This analysis is the loose Bayesian analogue of the frequentist post-hoc power analysis, which has been much criticised [e.g., @lakens_20_2014]. Most importantly, an assumption of the present analysis is that the data from @moffatt_inner_2020 is our best source of information regarding the main effect of interest (in addition to the prior we specified earlier). However, the present analysis also differs from the frequentist post-hoc power analysis on several grounds. First, with the present analysis, we do not aim to assess the ability of our statistical test to pass some dichotomic threshold (e.g., accept/reject). Instead, we aim to assess how the $\text{BF}_{10}$ (i.e., the evidence for the alternative hypothesis, relative to the null hypothesis) behaves with varying sample sizes. Second, the present analysis relies on the posterior predictive distribution of the model fitted on the data from @moffatt_inner_2020, which naturally incorporates uncertainty about the effect of interest. By simulating datasets of varying sample sizes from the posterior predictive distribution (and by relying on a large number of simulations), uncertainty about the effect size is naturally incorporated into the simulation.

## Manipulating rumination within-subject

In @nalborczyk_dissociating_2020, we manipulated the modality of rumination (whether it is verbal or non-verbal) in a between-subject manner to avoid order effects... In contrast to this approach, @moffatt_inner_2020 asked participants to ruminate and then distract themselves (or reciprocally), after an induced stressor (an induced failure)...

```{r order, fig.cap = "Average log-EMG amplitude by muscle and condition. The black dots and intervals represent the by-group average and 95\\% confidence interval (N = 26). The horizontal white line in the violin plot represents the median. The grey dots represent the individual-level average natural logarithm of the EMG amplitude by muscle and condition."}
# position for groups
pd <- position_dodge(0.8)

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order, labels = c("RUM -> DIS", "DIS -> RUM")) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value, colour = order, fill = order) ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun.y = mean,
        geom = "line", size = 1,
        aes(group = interaction(muscle, order) ),
        # aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun.y = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1) +
    theme(
      legend.title = element_blank(),
      # legend.position = c(0.85, 0.85),
      legend.position = "top",
      legend.background = element_rect(size = 0.5, linetype = "solid", colour = "black")
    )
```

About the order effects, @moffatt_inner_2020 say:

> "Unless otherwise reported, the inclusion of order in which the conditions were completed as a between-subjects variable as part of a mixed-design ANOVA produced no significant main effects or interactions involving order."

Unfortunately, the same line of reasoning applies for testing the effect of the order, which is even less powered than the test of the main effect of interest, rendering it practically uninformative...

## Does everyone?

@haaf_developing_2017...

```{r everyone, fig.cap = "Inter-individual variability in the main effect of interest (i.e., the difference between the rumination and distraction conditions). Green dots and lines represent the average natural logarithm of the EMG amplitude of participants that showed a higher EMG amplitude in the rumination condition than in the distraction condition, whereas orange dots and lines represent the average natural logarithm of the EMG amplitude of participants that showed a higher EMG amplitude in the distraction condition than in the rumination condition."}
# does everyone plot
df2 %>%
    pivot_longer(cols = FRO:OOI, names_to = "muscle") %>%
    pivot_wider(names_from = condition, values_from = value) %>%
    mutate(ruminator = ifelse(RUM > DIS, 1, 0) ) %>%
    mutate(ruminator = as.factor(ruminator) ) %>%
    mutate(rum_dis = RUM - DIS) %>%
    pivot_longer(cols = RUM:BAS, names_to = "condition") %>%
    ggplot(aes(x = condition, y = value, colour = ruminator) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.15,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_point(size = 2, alpha = 0.8, show.legend = FALSE) +
    # plotting individual trajectories
    geom_line(aes(group = ID), size = 0.5, alpha = 0.8, show.legend = FALSE) +
    # facetting by muscle
    facet_wrap(~muscle) +
    # axis labels
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    # scale_color_gradient2(midpoint = 0, low = "red", mid = "green", high = "blue", space = "Lab")
    scale_colour_brewer(palette = "Dark2", direction = -1) +
    scale_fill_brewer(palette = "Dark2", direction = -1)
```

Huge inter-individual variability... which leads to the next point, what is the relation between self-reports and EMG?

## Relation between self-report and EMG correlates

...

# Discussion and conclusions

...

# Supplementary materials {#supp}

Reproducible code and figures are available at https://github.com/lnalborczyk/inner_experience_EMG.

# Acknowledgements {-}

...

# References {-}

```{r create_r_references, cache = FALSE}
r_refs(file = "bib/r-references.bib")
```
